### 2025-02-18 paper.qmd
v1 of data description graphs

---
title: "Prediction Developing Countries' Life Expectancy"
author: 
  - Siddharth Gowda
  - Syed Hasan
  - Arshh Relan
date: today
date-format: long
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

\newpage

# Introduction

\newpage

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(knitr)
library(modelsummary)
library(janitor)
library(ggplot2)
library(knitr)
```

# Data Description

```{r}
#| include: false
#| warning: false
#| message: false

raw_data = read_csv("./Life_Expectancy_Data.csv")
clean_data = raw_data %>% clean_names() %>% drop_na()

```

```{r}
#| include: false
#| warning: false
#| message: false
data_interest = clean_data %>% filter(status == "Developing") %>% rename(gdp_capita = gdp) %>% select(life_expectancy, infant_deaths, hiv_aids, gdp_capita, schooling, percentage_expenditure, alcohol)
data_interest <- data_interest %>% 
  mutate(alcohol = case_when(alcohol <= 1.08 ~ "Negligible", 
                             TRUE ~ "Consumed"))

```

## Response Variable

```{r}
#| label: fig-hislfe
#| echo: false
#| warning: false
life_expectancy_stats <- data_interest %>%
  summarise(
    Mean = mean(life_expectancy, na.rm = TRUE),
    Median = median(life_expectancy, na.rm = TRUE),
    Std_Dev = sd(life_expectancy, na.rm = TRUE),
    Min = min(life_expectancy, na.rm = TRUE),
    Max = max(life_expectancy, na.rm = TRUE)
  )

ggplot(data_interest, aes(x = life_expectancy)) +
  geom_histogram(binwidth = 4, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Life Expectancy", x = "Life Expectancy", y = "Countries")

# kable(life_expectancy_stats, caption = "Basic Statistics for Life Expectancy")
```

\newpage

## Predictor Variables

```{r}
#| label: fig-predvars
#| layout-ncol: 3
#| layout-nrow: 2
#| fig-cap: ""
#| fig-subcap: ["", "", "", "", "", ""]
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false
#| fig.show: "hold"

ggplot(data_interest, aes(x = hiv_aids)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of HIV/AIDS", x = "HIV/AIDS Deaths per 1000 live births", y = "Countries")

ggplot(data_interest, aes(x = gdp_capita)) +
  geom_histogram(binwidth = 2500, fill = "lightgreen", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of GDP per Capita", x = "GDP per Capita", y = "Countries")

ggplot(data_interest, aes(x = infant_deaths)) +
  geom_histogram(binwidth = 50, fill = "lightcoral", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Infant Deaths", x = "Infant Deaths per 1000 People", y = "Countries")

ggplot(data_interest, aes(x = percentage_expenditure)) +
  geom_histogram(binwidth = 250, fill = "lightblue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Health Expenditure", x = "Health Expenditure (as a percentage of GDP per capita)", y = "Countries")

ggplot(data_interest, aes(x = schooling)) +
  geom_histogram(binwidth = 1, fill = "lightgoldenrod", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Schooling", x = "Average Number of years of Schooling", y = "Countries")

ggplot(data_interest, aes(x = alcohol, fill = alcohol)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Number of Countries that Consume Alcohol", 
       x = "Alcohol Category", y = "Number of Countries") +
  scale_fill_manual(values = c("Negligible" = "skyblue", "Consumed" = "salmon"))

```

```{r}
#| include: false
#| warning: false
#| message: false

alcohol_status <- data_interest %>%
  group_by(alcohol) %>%
  summarise(Count = n()) %>%
  as.data.frame()

# Display the table using kable
# kable(alcohol_status, caption = "Number of Countries with Negligible and Consumed Alcohol Data")

```

\newpage

# Preliminary Results